PeerMind

### 2016-08-16 paper.tex
Improved description of delegation.

% Load basic packages
\usepackage{balance}  % to better equalize the last page
\usepackage{graphics} % for EPS, load graphicx instead 
\usepackage[T1]{fontenc}
\usepackage{txfonts}
\usepackage{times}    % comment if you want LaTeX's default font
\usepackage[pdftex]{hyperref}
\usepackage{url}      % llt: nicely formatted URLs
\usepackage{color}
\usepackage{textcomp}
\usepackage{booktabs}
\usepackage{ccicons}
\usepackage{todonotes}
\usepackage{pgf}
\usepackage{tikz}
\usepackage{tkz-graph}
\usepackage{mathtools}

@@

\begin{itemize}
\item Member $M_0$ decides not to delegate at all.
\item Member $M_1$ delegates to $M_0$ and $M_3$ with ratios $0.75$ ($r_{1,0}$) and $0.25$ ($r_{1,3}$), respectively.
\item Member $M_2$ delegates to $M_3$ with ratio $1.00$ ($r_{2,3}$).
\item Member $M_3$ delegates to $M_1$, $M_2$, and $M_4$ with ratios $0.40$ ($r_{3,1}$), $0.40$ ($r_{3,2}$), and $0.20$ ($r_{3,4}$), respectively.
\item Member $M_4$ decides not to delegate.
\end{itemize}

The sum of ratios of all delegations of a member has to be $1.0$.
Members $M_0$ and $M_4$ decided to not participate in delegation and if they do not vote they are
counted towards non-voting population.

@@

    \Vertex[Math,L=M_0]{M0}
    \EA[Math,L=M_1](M0){M1}
    \SO[Math,L=M_3](M1){M3}
    \WE[Math,L=M_2](M3){M2}
    \EA[Math,L=M_4](M3){M4}

    \Edge[label=$0.40$](M1)(M3)
    \Edge[label=$0.75$](M0)(M1)
    \Edge[label=$0.25$](M3)(M1)
    \Edge[label=$0.20$](M4)(M3)

    \tikzset{EdgeStyle/.append style={bend right}}
    \Edge[label=$1.00$](M3)(M2)
    \Edge[label=$0.40$](M2)(M3)
  \end{tikzpicture}
  \caption{Example delegation network. Edges are directed towards a member who is delegating, representing how their
  vote is computed from delegates' votes. E.g., $V_{M_1} = 0.75 V_{M_0} + 0.25 V_{M_3}$}~\label{fig:delegation-network}
\end{figure}

We can represent these delegations as a graph we call \emph{delegation network} as seen in
Figure~\ref{fig:delegation-network}.
We direct edges towards a member who is delegating, to represent how their vote is computed from delegates' votes.
For example, member $M_1$ delegates the decision to members $M_0$ and $M_3$.
$M_0$ and $M_3$ become $M_1$'s delegates.
As a consequence, $M_1$'s vote $V_{M_1}$ is computed from $M_0$'s and $M_3$'s votes ($V_{M_0}$ and $V_{M_3}$).

@@

\begin{displaymath}
\begin{array}{rcl}
V_{M_0} & = & V_{M_0} \\
V_{M_1} & = & r_{1,0} V_{M_0} + r_{1,3} V_{M_3} = 0.75 V_{M_0} + 0.25 V_{M_3} \\
V_{M_2} & = & r_{2,3} V_{M_3} = 1.00 V_{M_3} \\
V_{M_3} & = & r_{3,1} V_{M_1} + r_{3,2} V_{M_2} + r_{3,4} V_{M_4} = \\
& = & 0.40 V_{M_1} + 0.40 V_{M_2} + 0.20 V_{M_4} \\
V_{M_4} & = & V_{M_4} \\
\end{array}
\end{displaymath}

When only some members vote ($\mathbf{v}$) we can use a delegation network to compute votes
of all members ($\mathbf{x}$).
We convert a delegation network to a matrix $\mathrm{D}$:

\begin{displaymath}
\mathrm{D} = \left[d_{ij}\right] = \begin{dcases*}
 1 & if $i = j$ \\
 -r_{i,j} & \parbox[t]{.55\columnwidth}{if $M_i$ did not vote, but delegates to $M_j$ with ratio $r_{i,j}$} \\
 0 & otherwise \\
\end{dcases*}
\end{displaymath}

We solve the matrix equation for $\mathbf{x_v}$:

\begin{displaymath}
\mathrm{D} \mathbf{x_v} = \mathbf{v},\quad \mathbf{v} = \left[v_{i}\right] = \begin{dcases*}
 V_{M_i} & if $M_i$ voted \\
 0 & otherwise \\
\end{dcases*}
\end{displaymath}

To continue with our example, let's say that $M_0$ and $M_2$ voted for a decision with $0.7$ and against with $0.4$,
respectively.
We are using score voting on $[-1, 1]$ range, so those votes are represented with $V_{M_0} = 0.7$ and
$V_{M_2} = -0.4$.

@@

\begin{displaymath}
\mathbf{v} = \left[ \begin{array}{c}
0.70 \\
0 \\
-0.4 \\
0 \\
0 \\
\end{array} \right],\quad \mathbf{x_v} = \left[ \begin{array}{c}
0.70 \\
0.54 \\
-0.40 \\
0.056 \\
0 \\
\end{array} \right]
\end{displaymath}

\todo[inline]{Max: FYI, computing who to discard from the population as ``non-voting'' is a separate computation in itself.
You need to figure out what votes can be calculated partially as a consequence of delegation, and which cannot.
Use indicator vectors denoting 1 if member has voted, else 0.
\\
Mitar: This is exactly what $\mathbf{w}$ is later on.}

There are few details we have to look into.
Despite delegation some votes or parts of votes can still happen to be discarded.
Members who did not vote might have not declared their delegates.
Other members could delegate to them as well, potentially propagating lack of a vote further.
In our example, member $M_4$ has not voted so their vote is discarded.
As a consequence $M_3$'s vote cannot be computed fully, which influences $M_1$'s vote as well.

In such case we want to normalize ratios of other delegates, those who voted or for which we can
compute their votes recursively.
So $r_{3,1}$ and $r_{3,2}$ are normalized from $0.40$ to $0.50$.

@@

Moreover, from $\mathbf{x_v}$ alone we cannot see which elements represent vote $0$ and which
represent that it was not possible to compute a vote for a member.

@@

\begin{displaymath}
\mathbf{w} = \left[w_{i}\right] = \begin{dcases*}
 1 & if $M_i$ voted \\
 0 & otherwise \\
\end{dcases*},\quad \mathbf{a} = \left[a_{i}\right] = \begin{dcases*}
 1 & if $M_i$ abstained \\
 0 & otherwise \\
\end{dcases*}
\end{displaymath}

We solve $\mathrm{D} \mathbf{x_w} = \mathbf{w}$ and $\mathrm{D} \mathbf{x_a} = \mathbf{a}$ for
$\mathbf{x_w}$ and $\mathbf{x_a}$, respectively.
Non-zero elements of $\mathbf{x_w}$ correspond to members who voted or for which we can compute
a vote.
We define a normalized matrix:

\begin{displaymath}
\mathrm{D'} = \left[d'_{ij}\right] = \begin{dcases*}
 1 & if $i = j$ \\
 0 & if $x_{w_j} = 0$ ($V_{M_j}$ cannot be computed) \\
 -\frac{r_{i,j}}{R_i} & \parbox[t]{.55\columnwidth}{if $M_i$ did not vote, but delegates to $M_j$ with normalized ratio $\frac{r_{i,j}}{R_i}$, $R_i = \sum_{x_{w_k} \ne 0} r_{i,k}$} \\
 0 & otherwise \\
\end{dcases*}
\end{displaymath}

To compute votes with normalized ratios for only those delegates who voted or for which we can compute votes recursively,
we solve $\mathrm{D'} \mathbf{x} = \mathbf{v}$ for $\mathbf{x}$:

\begin{displaymath}
\mathbf{x} = \left[ \begin{array}{c}
0.70 \\
0.54 \\
-0.40 \\
0.07 \\
0 \\
\end{array} \right]
\end{displaymath}

If an element of $\mathbf{x_w}$ is equal to $0$, then we could not compute a vote for that member and we count
it towards non-voting population.
We just have to make sure that they are not in fact abstentions.
We check corresponding elements of $\mathbf{x_a}$ and if they are non-zero we count them as abstentions,
otherwise as non-votes.

The result (average of scores) without delegation is $\overline{v} = 0.15$ and
with delegation is $\overline{x} = 0.23$.
Delegation pushes decision slightly in favor.

We pass computed votes and number of abstentions into computation of dynamic quorum to determine
a quorum and credibility, in the same manner as if members voted themselves.

>>>>

% Load basic packages
\usepackage{balance}  % to better equalize the last page
\usepackage{graphics} % for EPS, load graphicx instead 
\usepackage[T1]{fontenc}
\usepackage{txfonts}
\usepackage{times}    % comment if you want LaTeX's default font
\usepackage[pdftex]{hyperref}
\usepackage{url}      % llt: nicely formatted URLs
\usepackage{color}
\usepackage{textcomp}
\usepackage{booktabs}
\usepackage{ccicons}
\usepackage{todonotes}
\usepackage{pgf}
\usepackage{tikz}
\usepackage{tkz-graph}
\usepackage{mathtools}
\usepackage{amsmath}

@@

\begin{itemize}
\item Member $M_1$ decides not to delegate at all.
\item Member $M_2$ delegates to $M_1$ and $M_4$ with ratios $0.75$ ($r_{1,0}$) and $0.25$ ($r_{1,3}$), respectively.
\item Member $M_3$ delegates to $M_4$ with ratio $1.00$ ($r_{2,3}$).
\item Member $M_4$ delegates to $M_2$, $M_3$, and $M_5$ with ratios $0.40$ ($r_{3,1}$), $0.40$ ($r_{3,2}$), and $0.20$ ($r_{3,4}$), respectively.
\item Member $M_5$ decides not to delegate.
\end{itemize}

The sum of ratios of all delegations of a member has to be $1.0$.
Members $M_1$ and $M_5$ decided not to participate in delegation and if they do not vote they are
counted towards non-voting population.

@@

    \Vertex[Math,L=M_1]{M1}
    \EA[Math,L=M_2](M1){M2}
    \SO[Math,L=M_4](M2){M4}
    \WE[Math,L=M_3](M4){M3}
    \EA[Math,L=M_5](M4){M5}

    \Edge[label=$0.40$](M2)(M4)
    \Edge[label=$0.75$](M1)(M2)
    \Edge[label=$0.25$](M4)(M2)
    \Edge[label=$0.20$](M5)(M4)

    \tikzset{EdgeStyle/.append style={bend right}}
    \Edge[label=$1.00$](M4)(M3)
    \Edge[label=$0.40$](M3)(M4)
  \end{tikzpicture}
  \caption{Example delegation network. Edges are directed towards a member who is delegating, representing how their
  vote is computed from delegates' votes. E.g., $V_2 = 0.75 V_1 + 0.25 V_4$}~\label{fig:delegation-network}
\end{figure}

We can represent these delegations as a graph we call \emph{delegation network} as seen in
Figure~\ref{fig:delegation-network}.
We direct edges towards a member who is delegating, to represent how their vote is computed from delegates' votes.
For example, member $M_2$ delegates the decision to members $M_1$ and $M_4$.
$M_1$ and $M_4$ become $M_2$'s delegates.
As a consequence, $M_2$'s vote $V_2$ is computed from $M_1$'s and $M_4$'s computed votes ($V_1$ and $V_4$)
if $M_2$ does not vote themselves.

With $v_i$ we represent a vote of a member $M_i$ if they voted, otherwise it is $0$.
$V_i$ represents their vote, if they have voted, or a computed vote based on their delegations.

@@

\begin{displaymath}
\begin{array}{rcl}
V_1 & = & \begin{dcases*}
 v_1 & if $M_1$ votes \\
 0 & otherwise \\
\end{dcases*} \\
V_2 & = & \begin{dcases*}
 v_2 & if $M_2$ votes \\
 r_{1,0} V_1 + r_{1,3} V_4 = 0.75 V_1 + 0.25 V_4 & otherwise \\
\end{dcases*} \\
V_3 & = & \begin{dcases*}
 v_3 & if $M_3$ votes \\
 r_{2,3} V_4 = 1.00 V_4 & otherwise \\
\end{dcases*} \\
V_4 & = & \begin{dcases*}
 v_4 & if $M_4$ votes \\
 r_{3,1} V_2+ r_{3,2} V_3 + r_{3,4} V_5 = & otherwise \\
 = 0.40 V_2 + 0.40 V_3 + 0.20 V_5 & \\
\end{dcases*} \\
V_5 & = & \begin{dcases*}
 v_5 & if $M_5$ votes \\
 0 & otherwise \\
\end{dcases*} \\
\end{array}
\end{displaymath}

When only some members vote we can use a delegation network to compute votes
of all members.
We convert a delegation network to a matrix $D$:

\begin{displaymath}
D = \left[d_{ij}\right] = \begin{dcases*}
 1 & if $i = j$ \\
 -r_{i,j} & \parbox[t]{.55\columnwidth}{if $M_i$ did not vote, but delegates to $M_j$ with ratio $r_{i,j}$} \\
 0 & otherwise \\
\end{dcases*}
\end{displaymath}

We solve the matrix equation for $\boldsymbol{V}$:

\begin{displaymath}
\mathrm{D} \boldsymbol{V} = \boldsymbol{v},\quad \boldsymbol{v} = \left[v_{i}\right],\quad \boldsymbol{V} = \left[V_{i}\right]
\end{displaymath}

To continue with our example, let's say that $M_1$ and $M_3$ voted for a decision with $0.7$ and against with $0.4$,
respectively.
We are using score voting on $[-1, 1]$ range, so those votes are represented with $v_1 = V_1 = 0.7$ and
$v_3 = V_3 = -0.4$.

@@

\begin{displaymath}
\boldsymbol{v} = \left[ \begin{array}{c}
0.7 \\
0 \\
-0.4 \\
0 \\
0 \\
\end{array} \right],\quad \boldsymbol{V} = \left[ \begin{array}{c}
0.70 \\
0.54 \\
-0.40 \\
0.056 \\
0 \\
\end{array} \right]
\end{displaymath}

\todo[inline]{Max: FYI, computing who to discard from the population as ``non-voting'' is a separate computation in itself.
You need to figure out what votes can be calculated partially as a consequence of delegation, and which cannot.
Use indicator vectors denoting 1 if member has voted, else 0.
\\
Mitar: This is exactly what $\boldsymbol{w}$ is later on.}

There are few details we have to look into.
Despite delegation some votes or parts of votes can still happen to be discarded.
Members who did not vote might have not declared their delegates.
Other members could delegate to them as well, potentially propagating lack of a vote further.
In our example, member $M_5$ has not voted so their vote is discarded.
As a consequence $M_4$'s vote cannot be computed fully, which influences $M_2$'s vote as well.

@@

Moreover, from $\boldsymbol{V}$ alone we cannot see which elements represent vote $0$ and which
represent that it was not possible to compute a vote for a member.

@@

\begin{displaymath}
\boldsymbol{w} = \left[w_i\right] = \begin{dcases*}
 1 & if $M_i$ voted \\
 0 & otherwise \\
\end{dcases*},\quad \boldsymbol{a} = \left[a_i\right] = \begin{dcases*}
 1 & if $M_i$ abstained \\
 0 & otherwise \\
\end{dcases*}
\end{displaymath}

We solve $D \boldsymbol{W} = \boldsymbol{w}$ and $\mathrm{D} \boldsymbol{A} = \boldsymbol{a}$ for
$\boldsymbol{W} = \left[W_{i}\right]$ and $\boldsymbol{A} = \left[A_{i}\right]$, respectively.

Non-zero elements of $\boldsymbol{W}$ correspond to members who voted or for which we can compute
a vote.

\todo[inline]{If there is a clique in delegation, without any vote inside a clique, the matrix becomes singular.
We should find a way to detect such rows and set corresponding elements of $\boldsymbol{W}$ to $0$ (or the row
to identity row before we try to compute $\boldsymbol{W}$).
This is exactly what we want in fact, that those in such a clique are seen as non-computable.
Then the rest works.}

We define a normalized matrix:

\begin{displaymath}
D' = \left[d'_{ij}\right] = \begin{dcases*}
 1 & if $i = j$ \\
 0 & if $W_j = 0$ ($V_j$ cannot be computed) \\
 -r'_{i,j} & if $M_i$ did not vote \\
 0 & otherwise \\
\end{dcases*}
\end{displaymath}

Where:

\begin{displaymath}
r'_{i,j} = \frac{r_{i,j}}{R_i},\quad R_i = \sum_{W_k \ne 0} r_{i,k}
\end{displaymath}

We normalize ratios to only those delegates who voted or for which we can
compute their votes recursively.
So $r_{3,1}$ and $r_{3,2}$ are normalized from $0.40$ to $0.50$.

To compute votes with normalized ratios for only those delegates who voted or for which we can compute votes recursively,
we solve $D' \boldsymbol{V} = \boldsymbol{v}$ for $\boldsymbol{V}$:

\begin{displaymath}
\mathbf{V} = \left[ \begin{array}{c}
0.70 \\
0.54 \\
-0.40 \\
0.07 \\
0 \\
\end{array} \right]
\end{displaymath}

If an element of $\boldsymbol{W}$ is equal to $0$, then we could not compute a vote for the corresponding
member and we count it towards the non-voting population.
We just have to make sure that they are not in fact abstentions.
For elements $\boldsymbol{W} = 0$ we check their corresponding elements of $\boldsymbol{A}$ and if they are
non-zero we count them as abstentions, otherwise as non-votes.

\begin{itemize}
\item Number of directly voting members: $\left| v \right|$
\item Number of voting members: $\left| V \right| = \left| W_k \ne 0 \right|$
\item Number of abstentions: $\left| A \right| = \left| W_k = 0 \land A_k \ne 0 \right|$
\item Number of non-voting members: $\left| W_k = 0 \land A_k = 0 \right|$
\end{itemize}

The result (average of scores) without delegation is:

\begin{displaymath}
\overline{v} = \frac{\sum_{M_k voted} v_k}{\left| v \right|} = 0.15
\end{displaymath}

With delegation:

\begin{displaymath}
\overline{V} = \frac{\sum_{W_k \ne 0} V_k}{\left| V \right|} = 0.23
\end{displaymath}

In our example, delegation slightly pushes decision in favor.

We pass computed votes $V$ and number of abstentions $\left| A \right|$ into computation of dynamic quorum to determine
a quorum and credibility, in the same manner as if members voted themselves.

\todo[inline]{Do we want really to compute quorum using delegation, or should quorum be computed
using direct votes only?}
